Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-03-20 doc/leanappendix.tex
Fix capitalization
Clarified the unsound condition we discovered

\section{More information about the formally verified checker}
\label{appendix:lean}

We have implemented the rightmost two components of Figure~\ref{fig:chain}, namely the
proof checker and both model and weighted model counters, in the \lean{} programming language
and proof assistant [cite]. \lean{} implements a logical foundation in which expressions
have a computational interpretation, and just as with other proof assistants like Isabelle and Coq,
a function defined within the formal system can be compiled to efficient code.
At the same time, we can state and prove claims about the function within the system, thereby
verifying that the functions compute the intended results.

Verifying our checker and the model counters in this way serves at least two purposes.
When we run them on particular data sets, they provide high confidence
that the particular results we obtain are correct.
More broadly, however, our success also validates our general approach,
showing, in particular, that the CRAT proof format is sound.
Indeed, our formalization helped us discover a bug in our implementation
of the informal proof checker: in the ``\emph{s}'' step described at the end of
Subsection~\ref{subsection:syntax}, it is sound (and necessary) to allow defining
clauses for the new operations in the proof of $\obar{\lit}_1 \lor \obar{\lit}_2$,
but it is generally not sound to allow the input clauses or other added clauses.
[NOTE: let's add this specification to the end of Section 6.1.]
Our verified implementation of the checker tells us that the algorithm is sound and
provides a reference for other implementations.

@@

\subsection{Logical objects and operations}

@@

\subsection{Verification specifics}
\label{subsection:verification:specifics}

>>>>

\section{Detailed Information about the Formally Verified Checker}
\label{appendix:lean}

We have implemented the rightmost two components of Figure~\ref{fig:chain}, namely the
proof checker and both model and weighted model counters, in the \lean{} programming language
and proof assistant~\cite{demoura:cade:2021}. \lean{} implements a logical foundation in which expressions
have a computational interpretation, and just as with other proof assistants like Isabelle and Coq,
a function defined within the formal system can be compiled to efficient code.
At the same time, we can state and prove claims about the function within the system, thereby
verifying that the functions compute the intended results.

Verifying our checker and the model counters in this way serves at
least two purposes.  When we run them on particular data sets, they
provide high confidence that the particular results we obtain are
correct.  More broadly, however, our success also validates our
general approach, showing, in particular, that the CRAT proof format
is sound.  Indeed, our formalization helped us discover a bug in our
prototype proof checker: the declaration a sum operation via the
``\texttt{s}'' step (Section~\ref{subsection:syntax}) requires a hint
consisting of a sequence of clause Ids that provide a RUP proof of the
mutual exclusion condition $\obar{\lit}_1 \lor \obar{\lit}_2$.  It is
sound (and necessary) to allow these Ids to be those of defining
clauses for previous POG nodes, but, as we discovered, it is generally
not sound to allow them to be for input clauses or other added
clauses.  We therefore added the condition that only defining clauses
could be included in the hint, and we implemented a test of this rule
in our prototype checker.  Our verified implementation of the checker
tells us that the algorithm is sound and provides a reference for
other implementations.

@@

\subsection{Logical Objects and Operations}

@@

\subsection{Verification Specifics}
\label{subsection:verification:specifics}
